---
title: "Tidyverse and Baseball"
output: 
  html_notebook: 
    theme: spacelab
    toc: yes
---

Before doing anything, we'll load the 2 libraries we need: Lahman and tidyverse.

```{r}
library(Lahman)
library(tidyverse)
```

# Tibbles

Now that we have loaded the Lahman R library, we have access to all of the data frames contained there. However, the tidyverse has a new data structure called "tibbles" that act similarly to the data frames with less overhead.

One major difference can be seen with printing. Here I'll print the first 100 entries of the Batting dataframe. If I hadn't used the "head" command, all records would be printed to the notebook.

```{r}
head(Batting, 100)
```

With the tibble package, we get access to a new function called "glimpse", which I think does a better job of telling you what's in a table without overloading memory.

```{r}
batting <- as_tibble(Batting)
glimpse(batting)
```

# Dplyr

## filter

The first dplyr verb I'll demonstrate is filter. This is used to subset the rows of a tibble based on a condition. Here I'll limit our batting table to statistics that took place in the ["live ball era"](https://en.wikipedia.org/wiki/Live-ball_era), which started in 1920. This limits the number of entries to 83,196.

```{r}
glimpse(filter(batting, yearID >= 1920))
```

## select

```{r}
glimpse(Master)
```

```{r}
batting_names <- left_join(Batting, Master, by='playerID')
glimpse(batting_names)
```
```{r}
batting_names <- select(batting_names, nameGiven, AB, H, HR, RBI)
glimpse(batting_names)
```

## arrange

```{r}
glimpse(arrange(batting, desc(HR)))
```

## mutate

```{r}
glimpse(mutate(batting, AVG = H / AB))
```

# Magrittr

Magrittr uses something called "pipes" to string multiple functions together.

```{r}
best_hitters <- batting %>%
                left_join(Master, by='playerID') %>%
                mutate(AVG = H/AB) %>%
                mutate(FullName = paste(nameFirst, nameLast, sep=' ')) %>%
                arrange(desc(AVG)) %>%
                filter(yearID >= 1920, AB > 500) %>%
                select(FullName, AVG, HR, RBI)
best_hitters
```

